---
title: "Reproducible Research: Peer Assessment 1"
output:   html_document
keep_md: true
author: "Vish Murden"
date: "2015/08/17"
---

# Introduction

In this report we investigate the use of data as collected via an actvity monitoring device that captures personal movement.

The measurements are taken from a device that collects the number of steps taken over a 5 minute intervals through out the day. The data recording is taken over a 2 month period for a given anonymouse individual during the months of October - Novemeber 2012.

## Loading and preprocessing the data

The [data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) was originally sourced from the Coursera run Reproducible Reseach course, and a unzip copy retained in the data folder.

```{r readdata, message=FALSE}
library(dplyr)
library(lattice)

data <- read.csv("./data/activity.csv", colClasses=c("numeric", "character", "numeric"))
data$date <- as.Date(data$date, "%Y-%m-%d")
options(digits=2,scipen=999) # display doubles to 2dp, + disable scientific notation

```

Using this data, we'll ask a number of question and derive relavent meaning.

## What is mean total number of steps taken per day?

```{r, echo=FALSE}
filtered_data <- filter(data, data$steps !="NA")    # remove measurements where we have no data
groupbydate <- group_by(filtered_data, date)
```

Below we can see a summary of the total number of steps taken per day where a recording was available.

```{r}
data_summary <- summarise(groupbydate, totalsteps=sum(steps))
hist(data_summary$totalsteps, main = "Total steps taken per day", xlab = "Steps")
```


```{r calculatemeanmedian, echo=FALSE}
meansteps <-  mean(data_summary$totalsteps)
mediansteps <- median(data_summary$totalsteps)
```


It can be sean that the mean value of steps take is `r meansteps` whilst the median is `r mediansteps`

## What is the average daily activity pattern?

In the below plot we observe the average steps taken within the 5minute data within the day.

```{r calculateaveragedaily}
data_summary <- summarise(groupbydate, avgsteps=mean(steps))
plot(data_summary, type="l", main = "Average daily steps within 5min interval", ylab = "Steps", xlab = "Date")
```

```{r max5minsteps, echo=FALSE}
max5minterval <- filter(data_summary, avgsteps==max(data_summary$avgsteps))                        
```

A maximmum number of average steps of `r max5minterval$avgsteps` occured on `r max5minterval$date`

## Imputing missing values
```{r calcmissingdata, echo=FALSE}
missingData <- nrow(filter(data, is.na(data$steps)))
```
The original dataset comprises of `r missingData` samples in which a value was not provided, ie. NA. Such loss of data will skew our results a little.


```{r replaceNafunc, echo=FALSE}
replaceNa <- function(data, meansummarydata) {
    # replace all "NA" values with the lookup mean data for the date 
    # for which we have observed NA. If we have a mean value available
    nadata = filter(data, is.na(data$steps))
    for(i in 1:nrow(nadata)) {
        founditem <- filter(meansummarydata, date==nadata$date[i])
        if (nrow(founditem) > 0) {
            nadata$steps[i] = founditem$avgsteps[1]
        }
        else {
            nadata$steps[i] = 0
        }
            
    }
    
    nadata
}
```

```{r imputmissing}
reformmeddata <- merge(filtered_data, replaceNa(data,data_summary ), all=TRUE)
reformmeddata_summary <- summarise(group_by(reformmeddata,date), totalsteps=sum(steps))

# display mean value having replaced all NAs where possible
hist(reformmeddata_summary$totalsteps, main = "Total steps taken per day", xlab = "Steps")

reformmeddata_meansteps <-  mean(reformmeddata_summary$totalsteps)
reformmeddata_mediansteps <- median(reformmeddata_summary$totalsteps)
```

Having replaced where possible all NA values with corresponding mean values of the day, and 0 where we have a NA and no pre-existing mean observation for a date. We can observe overall mean number of steps now as `r reformmeddata_meansteps` and the median number of steps as `r reformmeddata_mediansteps`. It can be seen that then reformatted data seems to give slighly different values compared to earlier on, where the mean was `r meansteps` and median was `r mediansteps`. Notability the mean number of steps is lower.

## Are there differences in activity patterns between weekdays and weekends?
```{r createweekdayfactor, echo=FALSE}
# first need to figure what dates occured on Weekend or weekday, assign T/F
reformmeddata_summary <- summarise(group_by(reformmeddata,date), averagesteps=mean(steps))
weekday_reformmeddata_summary <- mutate(reformmeddata_summary, Weekday=weekdays(reformmeddata_summary$date, abbreviate = TRUE) %in% c("Mon", "Tue", "Wed", "Thu", "Fri"))

# now apply more relevant label instead of T/F
weekday_reformmeddata_summary$period <- factor(weekday_reformmeddata_summary$Weekday, levels = c(TRUE, FALSE), labels = c("Weekday", "Weekend"))
```

Below is a plot of average daily steps per day, broken out by average steps taken in the weekend and weekday.

```{r displayplot}
xyplot(averagesteps ~ date | period, data = weekday_reformmeddata_summary, layout = c(1,2), type = "l")
```


Looking at the plot, we cannot see any discernable difference between foot activities between the varied periods. 

