---
title: "Personal Movement Activity Moniting Report"
author: "Vish Murden"
date: "2015/08/17"
output: html_document
---

# Introduction

In this report we investigate the use of data as collected via an actvity monitoring device that captures personal movement. 

The measurements are taken from a device that collects the number of steps taken over a 5 minute intervals through out the day. The data recording is taken over a 2 month period for a given anonymouse individual during the months of October - Novemeber 2012.

The [data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) was originally sourced from the Coursera run Reproducible Reseach course, and a unzip copy retained in the data folder.

```{r readdata, message=FALSE}
library(dplyr)
data <- read.csv("./data/activity.csv", colClasses=c("numeric", "character", "numeric"))
data$date <- as.Date(data$date, "%Y-%m-%d")
options(digits=2,scipen=999) 'display doubles to 2dp, + disable scientific notation

```

Using this data, we'll ask a number of question and derive relavent meaning.

### What is mean total number of steps taken per day?

```{r, echo=FALSE}
filtered_data <- filter(data, data$steps !="NA")    # remove measurements where we have no data
groupbydate <- group_by(filtered_data, date)
data_summary <- summarise(groupbydate, totalsteps=sum(steps))
```

Below we can see a summary of the total number of steps taken per day where a recording was available.

```{r}
d <- as.Date(rep(data_summary$date,data_summary$totalsteps))
hist(data_summary$totalsteps, main = "Total steps taken per day", xlab = "Steps")
```


```{r calculatemeanmedian, echo=FALSE}
meansteps <-  mean(data_summary$totalsteps)
mediansteps <- median(data_summary$totalsteps)
```


It can be sean that the mean value of steps take is `r meansteps` whilst the median is `r mediansteps`
